---
title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India"
subtitle: 
author: 
  - "Joshua D. Merfeld <br> KDI School and IZA"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
                      fig.width = 9, fig.height = 3.5, fig.retina = 3,
                      out.width = "100%",
                      cache = FALSE,
                      echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      hiline = TRUE
                      )


library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')
# Double check
getwd()    # check


reportP <- function(pValue){
  if (pValue < 0.001){
    result <- "p < 0.001"
  } else {
    result <- sprintf("p = %.3f", pValue) # inserts a float into a string and simultaneously do rounding
  }
  return(result)
}

reportDec <- function(dec){
  
    result <- sprintf("%.3f", dec) # inserts a float into a string and simultaneously do rounding
  
  return(result)
}

reportDecOne <- function(dec){
  
    result <- sprintf("%.1f", dec) # inserts a float into a string and simultaneously do rounding
  
  return(result)
}

reportPercent <- function(dec){
  
    result <- sprintf("%.1f", 100*dec) # inserts a float into a string and simultaneously do rounding
  
  return(result)
}

commaNum <- function(large){
  
  
  result <- formatC(large, big.mark = ",")

  return(result)
}


# Shapefile of india
india_shape <- read_sf("data/spatial/shapefiles/state.shp")



# Coal plants
### Load raw data ----------------------------------------------------------------------------------

plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
# India only
plants <- plants %>% filter(Country=="India")
# Also want only things with a non-missing year
plants <- plants %>% filter(is.na(Year)==F)

# Just keep what we want 
plants <- plants %>% dplyr::select(plant_id = ParentID, 
                                   unit_id = Tracker.ID, 
                                   capacity = `Capacity.(MW)`, 
                                   year_built = Year, 
                                   year_retired = RETIRED, 
                                   lat = Latitude, 
                                   lon = Longitude)

plants <- plants %>% mutate(capacity = capacity/1000)

# long
plantsyearly <- c()
for (year in 1990:2015){
  plantstemp <- plants %>% filter(year_built<=year & (year_retired>year | is.na(year_retired)==TRUE))
  plantstemp <- plantstemp %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup() %>%
                  dplyr::select(plant_id, capacity, lat, lon)
  plantstemp$year <- year
  plantsyearly <- rbind(plantsyearly, plantstemp)
}


# Wind example
# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
# Go through two months worth of days
date <- as_date("2010-01-01")
villageswind <- c()
for (add_day in 1:100){
  wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-", month(date), "-", day(date), ".csv")) %>% as_tibble()
  wind <- wind[,1:2]
  colnames(wind) <- c("shrid", "indicator")
  villagestemp <- villages %>% left_join(wind, by = "shrid")
  villagestemp$date <- date
  
  villageswind <- rbind(villageswind, villagestemp)
  
  date <- date + 1
}

villageswind <- villageswind %>% st_set_crs(st_crs(districts))




# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yield1reducedform <- readRDS("pollution_development/draft/tables/yield1reducedform.rds")
yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
yield5heterogeneity <- readRDS("pollution_development/draft/tables/yield5heterogeneity.rds")
yield6leads <- readRDS("pollution_development/draft/tables/yield6leads.rds")
yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
labor2rural <- readRDS("pollution_development/draft/tables/labor2rural.rds")
labor3month <- readRDS("pollution_development/draft/tables/labor3month.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
opplot <- readRDS("pollution_development/draft/tables/opplot.rds")

```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
# style_duo_accent(
#   primary_color = "#006334",
#   secondary_color = "#A7A9AC",
#   inverse_header_color = "#FFFFFF"
# )
# style_mono_light(
#                  base_color = "#006334"
#                  )
style_duo(
          primary_color = "#A7A9AC", 
          secondary_color = "#006334"
          )
```

```{css, echo=F}
    /* Table width = 100% max-width */

    <!-- .remark-slide table{ -->
    <!--     width: 100%; -->
    <!-- } -->

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }
    
    tfoot {
      font-size: 80%;
    }
```





## Coal and energy worldwide

- Coal one of most common energy sources worldwide
  - Reached an all-time high in 2021 <small><small>(www.iea.org/reports/coal-fired-electricity)</small></small>

- india villages complain coal "power plant" (Adaniwatch)





---


```{r energypng, echo = FALSE, message = FALSE, warning = FALSE}

knitr::include_graphics(
                        path = "energy.png",
                        auto_pdf = TRUE
                        )

```





---

## Coal and energy in India

- From 1990 to 2010: coal generation from 42.4 gigawatts to more than 100
- Obvious benefits to more power <small><small>(Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017)</small></small>









---

## Downsides, too

- But power generation can have downsides, too
  - Especially true near plants: pollution a major concern
- Climate change and pollutants like SO2, mercury, etc. <small><small>(www.eia.gov/energyexplained/coal/coal-and-the-environment.php)</small></small>
- Long literature on negative effects of pollution 
  - Health <small><small>(Arceo, Hanna, and Oliva 2016; Brunekreef and Holgate 2002; Deryugina et al. 2019; Kampa and Castanas 2008; Pope III and Dockery 2006)</small></small>
  - Productivity/labor allocation <small><small>(Chang et al. 2019; Ebenstein, Lavy, and Roth 2016; Hanna and Oliva 2015; He, Liu, and Salvo 2019; Wen and Burke 2022)</small></small>




---

## This paper

- Effects of pollution from coal plants on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - Also instrument for particulate matter 2.5 (PM 2.5)



---

## This paper

- Effects on labor allocation decisions, as well
- Largest effects on labor and yield in same month
  - Around very beginning of the monsoon season (June)





---

## General idea
- Coal plants open in a given year
- Plot wind direction each day
  - Exposure means wind blowing from coal plant to village (within 30km)
  - Aggregate up to desired level (e.g. season, month, week)







---

## Data sources

```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

data <- matrix(NA, nrow = 8, ncol = 3)
rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
colnames(data) <- c("source", "geographic coverage", "temporal coverage")
data[1,1] <- "Asher et al. (2021)"
data[2,1] <- "Global Energy Monitor"
data[3,1] <- "NCAR"
data[4,1] <- "Hammer et al. (2020)"
data[5,1] <- "Angopadhyay et al. (2022)"
data[6,1] <- "Asher et al. (2021)"
data[7,1] <- "TerraClimate"
data[8,1] <- "National Sample Survey (NSS)"
data[1,2] <- "India"
data[2,2] <- "global"
data[3,2] <- "global"
data[4,2] <- "global"
data[5,2] <- "India"
data[6,2] <- "India"
data[7,2] <- "global"
data[8,2] <- "India"
data[1,3] <- " "
data[2,3] <- "yearly"
data[3,3] <- "daily"
data[4,3] <- "monthly"
data[5,3] <- "two seasons/year"
data[6,3] <- "yearly"
data[7,3] <- "monthly"
data[8,3] <- "daily interviews"

# Table
kable(
      data,
      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
      ) %>%
  footnote(
           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2,width = "4.5cm") %>%
  column_spec(c(3:4),width = "3.5cm") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_classic_2()

```






---

## Coal plants over time

```{r plants, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = india_shape, color = "gray85", fill = "gray80") +
    geom_point(data = plantsyearly, 
               aes(x = lon, y = lat, size = capacity),
               colour = "#006334", alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = year,
                      cumulative = FALSE
                      ) +
    labs(title = "year: {current_frame}") +
    theme(legend.position = "none")



```








---


## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r windbase, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = districts, color = "gray85", fill = "gray80") +
    geom_sf(data = villageswind %>% filter(date=="2010-01-01"),
            alpha = .5) +
    labs(title = "date: 2010-01-01") +
    theme(legend.position = "none")



```





---


## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r wind, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = villageswind, 
            aes(fill = indicator),
            alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = date,
                      cumulative = FALSE
                      ) +
    labs(title = "date: {current_frame}") +
    theme(legend.position = "none")



```





---

## Empirical strategy

First stage

$$pm_{it} = \alpha^1_{i} + \gamma^1_{t} + \phi^1 wind_{it} + \psi^1 weather_{it} + \varepsilon^1_{it}$$

Second stage

$$yield_{it} = \alpha^2_{i} + \gamma^2_{t} + \phi^2 \hat{pm}_{it} + \psi^2 weather_{it} + \varepsilon^2_{it}$$


- i indexes villages (districts for labor)
- t indexes time (where time differs based on the analysis)
- standard errors clustered at village (or district) throughout
- common use of stars (one indicates 0.10) throughout







---

## Where do coal plants open?


```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable,
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

```






---

## Reduced form - wind and yield


```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
kable(
      yield1reducedform,
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2.5cm") %>%
  row_spec(13, bold = TRUE) %>%
  row_spec(15, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```











---

## First stage - particulate matter and exposure, monthly level

```{r pollutiontable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable,
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

```










---

## PM and agricultural productivity

```{r yieldtablepm, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield2naive) <- c("(1)", "(2)", "(3)")
kable(
      yield2naive,
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2.5cm") %>%
  row_spec(5, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```


- Likely endogenous
  - Upward biased?











---

## Effects of PM on agricultural productivity

- We are really interested in the effect of _pollution_, not wind
- Use wind as instrument? Assumes...
  - Conditional on fixed effects and weather...
  - Wind only affects agriculture through pollution
- Reasonable?










---

## Effects of PM on agricultural productivity (IV)


```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
yield3ivmain <- yield3ivmain[,-1]
colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
rownames(yield3ivmain)[8] <- "F (first stage)"
kable(
      yield3ivmain,
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2cm") %>%
  row_spec(5, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

```




---

## Robustness checks

- Leads of wind direction
- Only post coal plant opening


---

## Heterogeneity

```{r heterogeneity, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
kable(
      yield5heterogeneity,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "4.5cm") %>%
  column_spec(c(2:6),width = "1.8cm") %>%
  row_spec(c(8), hline_after = TRUE) %>%
  row_spec(c(6), bold = TRUE) %>%
  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
  kable_classic_2()

```





---

## Taking stock

- We see clear negative effects of pollution on productivity
  - Absolute deviation of exposure is approx. 8 days, meaning 2-3 percent change in productivity
  - Using AD of PM, it's above 20 percent change
- Key question: what is the driver?
  - Land?
  - Labor?
- Hard to look explicitly at land, so let's look at labor


---


## National Sample Survey (NSS)

```{r nss, echo = FALSE, message = FALSE, warning = FALSE,}

laborplot

```


---

## Problem: only have district identifiers, so aggregate up


```{r aggregation, echo = FALSE, message = FALSE, warning = FALSE,}

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')

# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
wind <- wind %>% dplyr::select(shrid, days_sum)
villages <- villages %>% left_join(wind, by = "shrid")
villages <- villages %>% st_set_crs(st_crs(districts))

ggplot() + 
  theme_map() +
  geom_sf(data = districts, fill = NA) +
  geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
  scale_fill_viridis_c("exposure (days)", option = "viridis") +
  labs(x = "lon",
       y = "lat") +
  theme(legend.position = c(0.5, 0.05),
        legend.direction = "horizontal",
        legend.key.width = unit(0.3, 'cm'),
        legend.text = element_text(size = 8),
        legend.background=element_blank() )

```





---

## Wind direction and labor allocation, by rural status



```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labor1all) <- c("all", "all", "self", "wage", "farm", "non-farm")
colnames(labor2rural) <- c("all", "all", "self", "wage", "farm", "non-farm")

mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat1) <- c("panel a: all")
mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat2) <- c("panel b: rural")

labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])

kable(
      labortablerural,
      align = "cccccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:7),width = "1.5cm") %>%
  row_spec(c(5,10), hline_after = TRUE) %>%
  row_spec(c(1,6,12), bold = TRUE) %>%
  kable_classic_2()

```








---

## One last check

- Farm labor most common in July 
- Have monthly values of wind and pollution...
  - Heterogeneity by month of the season?
  - Monsoon season only
  






---

## Heterogeneity by month, labor



```{r labortablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labor3month) <- c("farm", "non-farm")
kable(
      labor3month,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(14), hline_after = TRUE) %>%
  row_spec(c(12), bold = TRUE) %>%
  kable_classic_2()

```






---

## Heterogeneity by month, yield



```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield4monthly) <- c("(1)", "(2)")
yield4monthly <- yield4monthly[c(1:14, nrow(yield4monthly)),]
kable(
      yield4monthly,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(12, bold = TRUE) %>%
  row_spec(14, hline_after = TRUE) %>%
  kable_classic_2()

```







---

## Coal plants or overall pollution?

- I use the location of coal plants to identify high-pollution areas
- However, some coal plants open part-way through sample
- Differences-in-differences combined with IV
  - Differences in coal pollution, specifically, and other forms of pollution





---

## Coal plants or overall pollution?


```{r ntltable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
kable(
      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2cm") %>%
  row_spec(c(7), bold = TRUE) %>%
  row_spec(9, hline_after = TRUE) %>%
  kable_classic_2()

```








---

## Conclusion

- Large growth in coal power in India
  - Other countries, as well
- Increase in electricity, but downsides as well
  - Pollution!
- We already have plenty of evidence on effects of pollution on health






---

## Conclusion

- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
  - Identification relies on within-village changes across seasons
- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
- Clear changes in labor allocation at beginning of the season
  - Yield only changes w.r.t. early-season pollution






---

## Conclusion

- Overall cost/benefit way beyond scope of paper
- Instead, results point to importance of the location of pollution sources
  - Downwind villages suffer
- 30km radius indicates relatively long-range effects






---

class: center, middle

<font size = "40"> Thank you! </font>

[https://joshmerfeld.github.io](https://joshmerfeld.github.io) <br>
Twitter: [@Josh\_Merfeld](twitter.com/Josh_Merfeld)








